---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# il.cbs.muni

<!-- badges: start -->
<!-- badges: end -->

il.cbs.muni is an analyst-oriented utility package for working with Israeli Central Bureau of Statistics (CBS) municipal data. It provides tools to handle the different quirks of CBS data formats, harmonize IDs across different years, and combine data points from multiple sources.

## Installation

You can install the released version of il.cbs.muni from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("il.cbs.muni")
```

Or install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("matanhakim/il.cbs.muni")
```

## Example

Let's say we would like to work with the [2021 yishuvim data](https://www.cbs.gov.il/he/publications/doclib/2019/ishuvim/bycode2021.xlsx) ('yishuvim' = points of residence) of the Israeli CBS. We will read the first two columns:

```{r}
library(il.cbs.muni)
library(readxl)
suppressMessages(library(dplyr))

# This is the specific path to the file in this package, but you can store it
# anywhere you want.
df_yishuvim <- read_excel("inst/extdata/bycode2021.xlsx", range = cell_cols("A:B")) |> 
  rename(yishuv_name = 1, yishuv_id = 2)
df_yishuvim
```

As you can see, the id column has two problems:

1. It's a numeric vector. This is problematic as our convention for an id throughout the il.verse is to have id's as characters.

2. Each yishuv id has a different length, ranging between 1-4. This is problematic since there is no difference between `67` and `0067`, and creates potential clashes  between some cities and regional councils when working with municipal data. For example, `31` is the **yishuv** id of Ofakim, but also the **municipal** id of Nahal Sorek regional council. Therefore, the convention is to use 4-character-long id for **every** yishuv.

using the `pad_yishuv_id()` function, we can fix this problem:

```{r}
df_yishuvim <- df_yishuvim |> 
  mutate(yishuv_id = pad_yishuv_id(yishuv_id))
df_yishuvim
```

Now every yishuv id is a string exactly 4 characters long.

We can verify this:

```{r}
df_yishuvim$yishuv_id |> 
  stringr::str_length() |> 
  range()
```

This solves a common problem when working with yishuvim data.
