---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# il.cbs.muni

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/il.cbs.muni)](https://CRAN.R-project.org/package=il.cbs.muni)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

il.cbs.muni is an analyst-oriented utility package for working with Israeli Central Bureau of Statistics (CBS) municipal data. It provides tools to read the CBS data files, handle the quirks of their formats, harmonize IDs across different years, and combine data points from multiple sources.

## Installation

You can install the released version of il.cbs.muni from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("il.cbs.muni")
```

Or install the development version from [GitHub](https://github.com/matanhakim/il.cbs.muni) with:

``` r
# install.packages("devtools")
devtools::install_github("matanhakim/il.cbs.muni")
```

## Reading municipal data

`read_cbs_muni()` reads a CBS municipal data file for a given year and data domain. It knows the per-year layout quirks - which sheet holds each domain, which rows form the (often multi-row) headers, and the aggregate summary rows that CBS added from 2022 onwards - so every returned row is a single municipality. Years 2003 through 2024 are supported.

```{r}
library(il.cbs.muni)
suppressMessages(library(dplyr))

df_muni <- read_cbs_muni(
  system.file("extdata", "p_libud_2021.xlsx", package = "il.cbs.muni"),
  year = 2021,
  data_domain = "physical"
)

df_muni |>
  select(1:4) |>
  glimpse()
```

## Reading index data

`read_cbs_index()` reads the socio-economic or peripherality index files. The CBS places the same geographic level in different table numbers across editions, so by default the function detects the level (municipality, locality or statistical area) from the file content and applies the matching header layout. When you already know the level, pass `unit_type` to override detection.

```{r}
df_index <- read_cbs_index(
  system.file("extdata", "24_22_375t1.xlsx", package = "il.cbs.muni"),
  year = 2019,
  index_type = "ses",
  quiet = TRUE
)

attr(df_index, "unit_type")

df_index |>
  select(1:4) |>
  glimpse()
```

## Harmonizing identifiers

CBS files use a numeric locality ID with a variable length, while the il.verse convention is a 4-character string. `pad_yishuv_id()` fixes that:

```{r}
df_yishuvim <- read_cbs_yishuv(
  system.file("extdata", "bycode2021.xlsx", package = "il.cbs.muni"),
  cols = c(1, 2)
) |>
  rename(yishuv_name = 1, yishuv_id = 2) |>
  mutate(yishuv_id = pad_yishuv_id(yishuv_id))

df_yishuvim
```

This matters because there is no difference between `67` and `0067` once stored as a number, which creates clashes between some cities and regional councils when working with municipal data. For example, `31` is the **yishuv** ID of Ofakim, but also the **municipal** ID of Nahal Sorek regional council, so the convention is a 4-character ID for **every** yishuv.

For regional councils, `modify_muni_id()` builds the municipal ID from the municipal status and locality ID, and its `rc_code` argument switches between the historical 2-digit code (e.g. `38`) and the 4-digit `5500 + code` form (e.g. `5538`) that CBS introduced in the 2024 municipal file.
