---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# il.cbs.muni

<!-- badges: start -->
<!-- badges: end -->

il.cbs.muni is an analyst oriented utility package to handle the different quirks of
  the Israeli CBS municipal data, harmonize id's and bring together data points
  from different years.

## Installation

You can install the development version of il.cbs.muni from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("matanhakim/il.cbs.muni")
```

## Example

Let's say we would like to work with the [2021 yishuvim data](https://www.cbs.gov.il/he/publications/doclib/2019/ishuvim/bycode2021.xlsx) ('yishuvim' = points of residence) of the Israeli CBS. We will read the first two columns:

```{r}
library(il.cbs.muni)
library(readxl)
suppressMessages(library(dplyr))

# This is the specific path to the file in this package, but you can store it
# anywhere you want.
df_yishuvim <- read_excel("inst/extdata/bycode2021.xlsx", range = cell_cols("A:B")) |> 
  rename(yishuv_name = 1, yishuv_id = 2)
df_yishuvim
```

As you can see, the id column has two problems:

1. It's a numeric vector. This is problematic as our convention for an id throughout the il.verse is to have id's as characters.

2. Each yishuv id has a different length, ranging between 1-4. This is problematic since there is no difference between `67` and `0067`, and creates potential clashes  between some cities and regional councils when working with municipal data. For example, `31` is the **yishuv** id of Ofakim, but also the **municipal** id of Nahal Sorek regional council. Therefore, the convention is to use 4-character-long id for **every** yishuv.

using the `pad_yishuv_id()` function, we can fix this problem:

```{r}
df_yishuvim <- df_yishuvim |> 
  mutate(yishuv_id = pad_yishuv_id(yishuv_id))
df_yishuvim
```

Now every yishuv id is a string exactly 4 characters long.

We can verify this:

```{r}
df_yishuvim$yishuv_id |> 
  stringr::str_length() |> 
  range()
```

This solves a common problem when working with yishuvim data.
